---
title: "Basic Simulation (Gaussian)"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=3)
library(ggplot2)
library(gridExtra)
```


Suppose that $X$ is a random variable with PDF $\pi(x)$. We are interested
in calculating $$\mathbb{E}_{\pi}(h(X))= \int h(x)\pi(x) \mathrm {dx}$$ This
might be an interval difficult to integrate and might not have a closed-form
solution, so we want to generate arbitrary points from the distribution
$\pi(x)$ to approximate this integral. There are cases where we only know $\pi(x)$ up to a multiplicative/normalizing constant. In these cases, importance sampling is used, which we will cover later. 

# rnorm

Suppose we know that $X \sim$ Gaussian(0,1). We draw randomly $n$
times from the standard Normal distribution using the function `rnorm`. 

```{r}
set.seed(21)
# 10 random draws from standard normal 
N10 <- rnorm(n = 10, mean = 0, sd = 1)
# 100 random draws from standard normal 
N100 <- rnorm(n = 100, mean = 0, sd = 1)
# 1000 random draws from standard normal 
N1000 <- rnorm(n = 1000, mean = 0, sd = 1)
# plot histogram of x_sim, follows a Gaussian(0,1)
n10 <- qplot(N10, geom="histogram", bins = 10)
n100 <- qplot(N100, geom="histogram", bins = 30)
n1000 <- qplot(N1000, geom="histogram", bins = 30)
grid.arrange(n10, n100, n1000, ncol = 3)
```


# dnorm

`dnorm` returns the probability density for the Gaussian distribution given parameters `x`, $\mu$, and $\sigma$, where Gaussian density function is $$f(x|\mu, \sigma) = \frac{1}{\sqrt{2\pi}\sigma} \exp \big\{-\frac{(x-\mu)^2}{2\sigma^2} \big\} $$

```{r}
# create a vector of z-scores
x_seq <- seq(-3,3,length.out = 100)
# find normal densities at each x
f_x_seq <- dnorm(x_seq)
# plot 
qplot(x = x_seq, y = f_x_seq, geom = "line", 
      xlab = "x", ylab = "density", main = "PDF of Gaussian(0,1)")
```

# pnorm 

`pnorm` returns the value of $F(x) = \mathbb P(X \leq x)$ at each $x, which is equivalent to 

$$
\int_{-\infty}^{x} \frac{1}{\sqrt{2\pi}\sigma} \exp \big\{-\frac{(x-\mu)^2}{2\sigma^2} \big\} dx
$$


```{r}
pvalues <- pnorm(x_seq)
plt1 = qplot(x = x_seq, y = pvalues, geom = "line",
             xlab = "x", ylab = "density", main = "CDF of Gaussian(0,1)") + 
  xlim(c(-3,3))
plt1
```

# qnorm

`qnorm` is inverse of `pnorm`. It returns $x$ from a given $F(x)$. 

```{r}
quantiles <- seq(0, 1, by = .01)
qvalues <- qnorm(quantiles)
plt2 <- qplot(x = qvalues, y = quantiles, geom = "line",
              xlab = "x", ylab = "density", main = "CDF of Gaussian(0,1)") + 
  xlim(c(-3,3))
plt2
```

Lastly, consider the function $h(x) = 10 e^{-2\|x-5\|}$. We want to find
$\mathbb{E}(h(X))$.

```{r}
h <- function(x) 10*exp(-2*abs(x-5))
# calculate h(x) at each of the random points drawn from standard normal and then approximate E(h(X))
mean(h(N1000))
```